### 2017-05-24 tex/methodology.tex

% Programs: `time`, bad; `\time -v`, better; oprofile(1); gprof(1); perf; Valgrind's
% cachegrind?; ...

% Performance Analysis; Run Time Analysis; Timing Methodology; Profiling Pitfalls; Run
% Time Measuring; Benchmarking Methodology; Performance Monitoring
\section{[TODO] Profiling Methodology}
\label{app:meth}

Plenty of things can go wrong when
profiling.
% measuring \alts{execution, run} times.
These are the ones I was aware of and tried to account for while doing measurements for
this \article.
\begin{itemize}
   \item \emph{CPU frequency scaling and boosting}.
      CPU frequencies are usually dynamic these days and automatically adjusted based on
      at least workload and temperature.  This should be disabled when measuring execution
      times.
      \begin{minted}[autogobble]{text}
         # TODO
      \end{minted}
      % See [5].  Not a problem when measuring cycles.
   \item \emph{Interrupts and context switches.}  The process being timed has to share
      resources with other processes and the operating system.  A context switch will not
      only increase wall time, it will also increase cycles because of TLB flushes and
      cache evictions. % See <https://en.wikipedia.org/wiki/Context_switch>.

      Processor shielding, % See [5].
      taking the minimum of several measurements, and assigning a very high priority can
      \alts{help here, mitigate these problems}.
      % This is likely better than using nice(1).
      \begin{minted}[autogobble]{text}
         # chrt -f 99 \time -v ./program
      \end{minted}
   \item \emph{CPU jumping.}
      % See <https://youtu.be/vrfYLlR8X8k?t=33m34s>.
      CPU pinning (setting processor/thread affinity) may help. % See [2] and [5].
   \item \emph{Cache warmth.}
      % See <https://youtu.be/vrfYLlR8X8k?t=41m00s>.
      When comparing different solutions on the same data set, later ones may benefit from 
      data having been loaded into cache already.  Checking whether swapping the order of
      measurements changes the results is a good idea.
\end{itemize}
These effects may \alts{add noise to, pollute} the results, render them irreproducible, or
invalidate them completely.

General mitigation/alleviation strategy: take the minimum execution time of \alts{all, a
number of} runs; all noise is additive (TODO: not the one caused by the cache being hot).
% See [7].

\subsection{Measuring CPU cycles}
\label{app:cycles}

Frequency scaling isn't a problem here.  I used the \mintinline{text}{ocount(1)} event
counting tool added to the OProfile (TODO: cite) project in version 0.9.9.  [TODO] This
uses \emph{hardware performance counters}.
% `perf(1)` seems to be a newer alternative to OProfile [8].

\begin{minted}[autogobble]{text}
   ophelp
   sudo chrt -f 99 ocount -e CPU_CLK_UNHALTED program
\end{minted}

TODO~\cite[78\psqq]{drepper2007}.

% [1]:  https://en.wikipedia.org/wiki/Profiling_(computer_programming)
% [2]:  https://en.wikipedia.org/wiki/Processor_affinity
% [3]:  https://en.wikipedia.org/wiki/OProfile
% [4]:  https://en.wikipedia.org/wiki/Perf_(Linux)
% [5]:  https://github.com/JuliaCI/BenchmarkTools.jl/blob/master/doc/linuxtips.md
% [6]:  http://stackoverflow.com/q/9006596 "On `/usr/bin/time` for benchmarks and `perf`"
% [7]:  https://youtu.be/vrfYLlR8X8k "Andrei Alexandrescu - Writing Fast Code"
% [8]:  http://oprofile-list.sf.narkive.com/rVbJlHdX/oprofile-vs-perf
% [9]:  http://oprofile.sourceforge.net/about/
% [10]: https://en.wikipedia.org/wiki/Hardware_performance_counter
% [11]: http://oprofile.sourceforge.net/doc/index.html

% vim: tw=90 sts=-1 sw=3 et fdm=marker