### 2017-05-26 tex/cpu-info.tex
Minor update

\section{Reading Information About the CPU}
\label{app:cpuinfo}
% https://stackoverflow.com/q/7281699
% https://unix.stackexchange.com/q/167038
% https://superuser.com/q/55776
There are many ways to display information about the processor(s) the operating system is
running on.  Among others, the \mintinline{text}{lscpu(1)} and
\mintinline{text}{getconf(1)} programs and the \mintinline{text}{/proc/cpuinfo}
pseudo-file on Linux.  This is how I checked my CPU's cache sizes, for example:
\begin{minted}[gobble=3]{bash}
   $ lscpu | grep 'L1d\|L2'
   L1d cache:             32K
   L2 cache:              512K
\end{minted}
%stopzone
This is what I used to get the cache line sizes:
\begin{minted}[gobble=3]{bash}
   $ getconf LEVEL1_DCACHE_LINESIZE; getconf LEVEL2_CACHE_LINESIZE
   64
   64
\end{minted}
%stopzone

% vim: tw=90 sts=-1 sw=3 et fdm=marker