### 2017-05-28 tex/basic-concepts.tex
Split basic-concepts.tex subsections into files

\section{Basic Concepts}
% \section{Basic Principles}
% \section{Key Terms} % That sounds boring.
Some architectural properties of hardware caches lead to important concepts for using them
effectively.

\input{tex/basic-concepts/lines.tex}
\input{tex/basic-concepts/prefetching.tex}
\input{tex/basic-concepts/assoc.tex}
\subsection{Cache Hits and Misses}
% TODO: what happens when we miss?  How bad is it?  What about \gls{smt}?
\input{tex/basic-concepts/spat_loc.tex}
\input{tex/basic-concepts/temp_loc.tex}
\subsection{Memory Access Pattern}
\label{sec:map}

% vim: tw=90 sts=-1 sw=3 et fdm=marker