### 2017-05-28 tex/basic-concepts/prefetching.tex

\subsection{Prefetching}
\label{sec:prefetch}
% http://ithare.com/c-for-games-performance-allocations-and-data-locality/
% http://ithare.com/c-for-games-performance-allocations-and-data-locality/2/
% https://en.wikipedia.org/wiki/Cache_prefetching

\emph{Cache prefetching} is a technique by which CPUs recognize predictable access
patterns (for instance sequential access to an array)... TODO

% vim: tw=90 sts=-1 sw=3 et fdm=marker