### 2017-05-30 tex/basic-concepts/prefetching.tex
Add new measurement; revise, expand, and tweak

\subsection{Prefetching}
\label{sec:prefetch}

% \cref{lst:TODO} is
Consider
a simplified version of \cref{lst:access-times} that, instead of using random accesses,
simply walks over the array sequentially.  It still follows the pointers to do this, but
the array is no longer shuffled.

% \emph{Cache prefetching} is a technique by which CPUs recognize predictable access
% patterns (for instance sequential access to an array)... TODO

% [1]: http://ithare.com/c-for-games-performance-allocations-and-data-locality/
% [2]: http://ithare.com/c-for-games-performance-allocations-and-data-locality/2/
% [3]: https://en.wikipedia.org/wiki/Cache_prefetching

% vim: tw=90 sts=-1 sw=3 et fdm=marker