### 2017-06-07 tex/list-vs-vector.tex
Add program comparing std::list and std::vector

\section{Example: \texttt{std::vector} vs. \texttt{std::list}}

\alts{%
   The C++ program shown in \cref{lst:ithare},
   The following C++ program,
}
\x{, adapted from \textcite{bigos},} initializes a number of \acrshort{stl} containers
with random numbers and measures the processor time needed to \alts{sum all of them,
compute the sum}.
%
I first ran it with \mintinline{cpp}{Container} being a type alias for
\mintinline{cpp}{std::list}, then for \mintinline{cpp}{std::vector}.%
\footnote{%
   Each compiled with \x{the \gls{gcc}} and \mintinline{text}{-O3} and
   \mintinline{text}{-march=native}
}

\begin{center}
   \inputminted[firstline=20]{cpp}{ithare/list-vs-vector.cpp}
   \captionof{listing}{Compute the Sum of a Container, adapted from \textcite{bigos}}
   \label{lst:ithare}
\end{center}

\newread\file
\openin\file=ithare/speedup.txt
\read\file to \speedup
\closein\file

% This doesn't work: `\num{\input{ithare/speedup.txt}}`.

When using \mintinline{cpp}{std::vector}, computing the sum is
\num[round-mode=places, round-precision=0]{\speedup} times faster.

% vim: tw=90 sts=-1 sw=3 et fdm=marker