### 2017-06-13 tex/com.tex

\subsection{Cache-Oblivious Model} % Cache-Oblivious Algorithms

% The \gls{com} is like the \gls{emm}, but we pretend not to know the values of B and M.
% This seems a bit silly.

% vim: tw=90 sts=-1 sw=3 et fdm=marker