### 2017-06-27 tex/graphics/xpose-speedup-plot.tex

\documentclass[tikz, border=1pt]{standalone}

\input{tex/datavis}

% FIXME: DRY.
\usepackage[binary-units]{siunitx} % Load `\kibi` and `\mebi` etc. (`binary-units`).

% FIXME: DRY.
\pgfkeys{%
   /pgf/number format/.cd,
   1000 sep={\,},
   min exponent for 1000 sep=4,
}

\begin{document}
\tikz \datavisualization[scientific axes=clean, x axis={
   logarithmic, label={Matrix Size (\si{\mebi\byte})}, length=0.8\textwidth, ticks={
         major at={1, 2, 4, 8, 16, 32, 64, 128, 256, 512, 1024},
      },
   }, y axis={
      include value=0, grid=at ticks, label={Speedup}, length=6cm,
   % ticks={major also at={1}}
   },
   visualize as scatter,
   scatter={style={mark=*, mark options={scale=.65}}}]
   data [read from file=xpose/speedup.csv, separator=\space];
\end{document}

% vim: ft=tex tw=90 sts=-1 sw=3 et fdm=marker